---
title: "R Notebook"
output: html_notebook
---


```{r}
rm(list=ls())
setwd("/home/tauro/Projects/quizziz")

library(jsonlite)
library(lubridate)
library(dplyr)
library(ggplot2)
#library(reshape2)
#library(purrr)

options(scipen=999)
```

```{r}
data <- stream_in(file("data_role_assignment.json"))
```

```{r}
#Some data preprocessing
params <- lapply(data$params, "[[", 2)
params <- setNames(do.call(rbind.data.frame, params), 
                       c("sessionId", "page", "userId"))

data$params <- NULL

data <- cbind(data, params)

rm(params)


data$serverTime <- ymd_hms(data$serverTime, tz="UTC")

data$day <- day(data$serverTime) 
data$month <- month(data$serverTime)
data$hour <- hour(data$serverTime)
data$wday <- wday(data$serverTime, label=TRUE)

categorical_cols <- c("version", "eventId", "sessionId", "page", "userId", "day",
                      "month", "hour", "wday")

data$wday <- as.factor(as.character(data$wday))

data[categorical_cols] <- lapply(data[categorical_cols], factor)

#Filling in missing userIds for the same session
data <- data %>% group_by(sessionId) %>% mutate("userId" = first(userId))

data$is_reg <- as.factor(as.integer(is.na(data$userId)))
```
  
<br />

The first metric we look at is the number of events. Here, *n_events* is the number of events for a particular session. 
```{r}
#Number of events per session across both versions
eventsPerSessionAB <- data %>% select(everything()) %>% 
  group_by(sessionId, version, is_reg) %>% summarise("n_events" = length(eventId)) %>% 
  arrange(desc(n_events)) 

eventsPerSessionAB
```
<br />
  
Testing for normality of *n_events* through a random sample (Shapiro's test)
```{r}
shapiro.test(sample(eventsPerSessionAB$n_events, 3000))
#Cannot assume normality, hence not ideal to perform t-test directly

```
  
  
  
<br />
The results clearly show conclusive evidence against the null hypothesis (that *n_events* is normally distributed)
  
  
Usually, one performs a two-sample Student's t-test of the means of two populations. But since the t-test assumes the data to be normally distributed and our data indicates otherwise, we need an alternative. 
  
  
We perform the Mann-Whitney-Wilcox test which does not assume normality.

```{r}
wilcox.test(n_events ~ version, data=eventsPerSessionAB)
#The number of events for both versions are not identically distributed

wilcox.test(n_events ~ is_reg, data=eventsPerSessionAB)
#The number of events for registered and non_registered sessions are not identically distributed


```
  
<br />

Based on the results of the Mann-Whitney-Wilcox test, we can conclude that the differences in *n_events* by version and registered status are statistically significant
```{r}

```



